---
title: "NE Greenland Trait Analysis: Setting the model in HMSC"
author: "Phoebe Armitage"
date: '2023-June'
output:
 html_document:
  toc: TRUE
  anchor_sections: TRUE
  code_download: TRUE 
  code_folding: "hide"
  fig_caption: TRUE
  theme:
    bootswatch: "flatly"
  bibliography: references.bib
  link-citations: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.show = "hide", results = "hide", eval = FALSE)
# to cite specific packages
knitr::write_bib(c("base", "bookdown", "rmarkdown"), "packages.bib")
```
## Load libraries 
```{r}
library(Hmsc); library(corrplot); library(vioplot); library(colorspace); library(parallel)
```
## Load data 
```{r}
load(file = "allData_A.R")
localDir = "."
#data.directory = file.path(localDir, "data")
model.directory = file.path(localDir, "models")
```

## The HMSC model
First we need to set up all the parameters and syntax.
### Environmental covariates (fixed effects)
Here we define the regression model of the fixed effects (XFormula) including the environmental covariates.
```{r}
XFormula <- ~ Depth + B.Temp + B.PSU + Oxygen + Turbidity  + Flourescence
```
### Traits 
Here we define the regression model that considers species traits
```{r}
TrFormula <- ~ Small + Small_med + Medium_S + Med_large + Large +
Globulose + Vermiform + Dorso_com + Later_com +  Upright +
  Calcareous + Siliceous + Chitinous + Cuticle + None_SK + 
  Asexual + Sexual_ext + Sexual_int + Sexual_brood + Pel_Plankto +
  Pel_Leci + Benthic_Dir + Sessile + Burrower + Crawler +
  Swimmer+ none_MV + Low + medium_MV + High +
  Deposit + Filter_Sus + Op_Sc + Pred + Arctic +
  Boreal + Cosmo
```
### Study design (random effects)
Now we create the study design from which we assign the random effect levels.
```{r}
studyDesign = data.frame(sample = as.factor(1:nrow(S)), habitat = as.factor(S$Habitat)) 
rL = HmscRandomLevel(units = studyDesign$sample) 
rL2 = HmscRandomLevel(units = studyDesign$habitat)
```
### Presence-absence data 
```{r}
Ypa <- 1*(Y>0) # covert species abundance matrix to presence absence records
```

### Defining the model
```{r}
m1 <- Hmsc(Y=Ypa, XData = X, XFormula = XFormula,
           TrData = Tr, TrFormula = TrFormula,
           phyloTree = P,
           studyDesign = studyDesign,
           ranLevels = {list("sample" = rL, "habitat" = rL2)}, 
           distr = "probit")
```

```{r}
#Reassign into syntax 
#NB: This is a good when running multiple models
models <- list(m1)
modelnames <- "presence_absence"
```

### Set MCMC parameters and run model
```{r}
samples_list = 250
thin_list = 1000
nChains = 4
for(Lst in 1:length(samples_list)){
  thin = thin_list[Lst]
  samples = samples_list[Lst]
  
  print(paste0("thin = ", as.character(thin),"; samples = ", as.character(samples)))
  nm = length(models)
  
  for(model in 1:nm) {
    #model = 1
    print(paste0("model = ",modelnames[model]))
    m = models[[model]]
    m = sampleMcmc(m, samples = samples, thin = thin, 
                   adaptNf=rep(ceiling(0.4*samples*thin),m$nr),
                   transient = ceiling(0.5*samples*thin), 
                   nChains = nChains, nParallel = nChains)
    models[[model]] = m
  }
  filename = paste("models/models_thin_", as.character(thin),
                   "_samples_", as.character(samples),
                   "_chains_",as.character(nChains),
                   ".Rdata",sep = "")
  save(models,modelnames,file=filename)
}
```

### Check model
```{r}
#Call model 
model <- models[[j]]
model$studyDesign
model$ranLevelsUsed
model$TrFormula
model$XFormula
model$samples
model$thin
model$transient
model$call
model$XData
```




