---
title: "NE Greenland Trait Analysis: Biodiversity and Functional Indices"
author: "Phoebe Armitage"
date: '2023-June'
output:
 html_document:
  toc: TRUE
  anchor_sections: TRUE
  code_download: TRUE 
  code_folding: "hide"
  fig_caption: TRUE
  theme:
    bootswatch: "flatly"
  bibliography: references.bib
  link-citations: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.show = "hide", results = "hide")
# to cite specific packages
knitr::write_bib(c("base", "bookdown", "rmarkdown"), "packages.bib")
```

## Libraries and data

To begin, we need to download our data for the analysis as well as
install any packages or libraries from R.
Note that for the first part (calculating the indices) it is the same code for both the abundance and biomass data sets. Therefore, you will need to run the code once for the abundance data set, and then run it again for biomass data set. You will see the lines of code that have a hash tag (#) that will need to be switched for abundance or biomass analyses. From the statistics section onwards you will not have to do this. 

### Libraries

```{r}
# load libraries 
library(vegan)     
library(devtools)  
library(ggplot2)
library(FD)
library(adiv)
library(RColorBrewer)
```

### Data

```{r}
# load taxa data 
load("allData_A.R")
# load("allData_B.R")

# assign new names
Phy <- P
Species <- Y
Traits <- Tr
Env <- X
S
```

## Biodiversity and Functional Indices
### Species diversity indices
The following code chunk is for species richness, total number of individuals, and the shannon and simpson diversity indices. See package vegand for more information. 

```{r}
Sp.index <- data.frame(Station = S$Station,
                       Habitat = S$Habitat,
                       Sp.rich = specnumber(Y), Total.ind = rowSums(Y),
                       Shannon = diversity(Y, index = "shannon"),
                      Simpson = diversity(Y, index = "simpson"))
Sp.index
```

#### Table with biodiversity indices
An example of how the new data looks like. Can use this code again to see the functional indices below. 

```{r}
knitr::kable(Sp.index,
 caption = "Biodiversity Indices. Metrics calculated using Vegan package in R Software")
```

### Functional Indices 
The code chunk below is for distance-based functional diversity indices and includes functional richness, funtional evennes, functional divergence, funtional dispersion, and Rao's quadratic entropy. CWM stands for community-level weighted means of trait values. See R package 'FD' for more information. 
```{r}
ex3<-dbFD(Tr, Y, corr="sqrt", m=4) 
ex3  

data<-data.frame(FRic = ex3$FRic, 
                 FEve = ex3$FEve, FDiv = ex3$FDiv, FDis =ex3$FDis,
                 RaoQ =ex3$RaoQ, Station = S$Station)

Sp.index_A <- merge(Sp.index, data, by = "Station")
#Sp.index_B <- merge(Sp.index, data, by = "Station")

CWM <- ex3$CWM
#write.csv(ex3$CWM,file = file.path(indices.directory, 'CWM_A.csv'))
#write.csv(ex3$CWM,file = file.path(indices.directory, 'CWM_B.csv'))

```

### Functional Redundancy 
The next code chunk is for functional redundancy and is from the R package Adiv. 
First we have to make a distnace matrix, and then follow it with the uniqueness function. This then produces functional redundancy at the community level and vulnerability at the species level. As vulnerability is calculated at the species level, the mean and standard deviation is then calculated for each community. 
```{r}
#make distance matrix
dist.MS <- function (Species, diag = FALSE, upper = FALSE, tol = 1e-07) 
{
    
    df <- data.frame(Species)
    if (!inherits(df, "data.frame")) 
        stop("df is not a data.frame")
    nlig <- nrow(df)
    d <- matrix(0, nlig, nlig)
    d.names <- row.names(df)
    fun1 <- function(x) {
        sum(abs(df[x[1], ] - df[x[2], ]))/sum(apply(cbind.data.frame(df[x[1], ], df[x[2], ]), 1, max))
    }
    df <- as.matrix(df)
    index <- cbind(col(d)[col(d) < row(d)], row(d)[col(d) < row(d)])
    
    d <- unlist(apply(index, 1, fun1))
    
    attr(d, "Size") <- nlig
    attr(d, "Labels") <- d.names
    attr(d, "Diag") <- diag
    attr(d, "Upper") <- upper
    attr(d, "method") <- "Marczewski–Steinhaus"
    attr(d, "call") <- match.call()
    class(d) <- "dist"
    return(d)
}

Dis <- dist.MS(Traits)
```

```{r}
Uni <- uniqueness(Species, Dis, abundance = TRUE)
Uni$red$Habitat <- S$Habitat
Uni$red

#And then Vulnerability
Vulnerability <- data.frame(Mean = colMeans(Uni$V, na.rm = T),
                             SD = apply(Uni$V, 2, sd, na.rm=TRUE))
boxplot(Uni$V)
Vulnerability

Sp.index_A$Vul <- Vulnerability$Mean
Sp.index_A$Red <- Uni$red$R

#Sp.index_B$Vul <- Vulnerability$Mean
#Sp.index_B$Red <- Uni$red$R

#write.csv(Sp.index_A, file = file.path(indices.directory, "spIndices_A.csv"))
#write.csv(Sp.index_B, file = file.path(indices.directory, "spIndices_B.csv"))
```
## Statistics 
### Read in both abundance and biomass data files. 
```{r}
Sp.index_A <- read.csv(file ="indices/spIndices_A.csv", row.names = 1)
Sp.index_B <- read.csv(file ="indices/spIndices_B.csv", row.names = 1)

Sp.index_A$Data <- 'Abundance' 
Sp.index_B$Data <- "Biomass"
All_indices <- rbind(Sp.index_A, Sp.index_B)
All_indices$Habitat <- as.factor(All_indices$Habitat)
```
Organise the data and find the means
```{r}
Fjord_indices<-subset(All_indices,Habitat=="Fjord")
Shelf_indices<-subset(All_indices,Habitat=="Shelf")
Shelfbreak_indices<-subset(All_indices,Habitat=="Shelfbreak")
Slope_indices<-subset(All_indices,Habitat=="Slope")

means_fjord <- apply(Fjord_indices[,c(3,5,6,7,8,9,11,12, 13, 14)], 2, mean)
means_shelf <- apply(Shelf_indices[,c(3,5,6,7,8,9,11,12, 13, 14)], 2, mean)
means_shelfbreak <- apply(Shelfbreak_indices[,c(3,5,6,7,8,9,11,12, 13, 14)], 2, mean)
means_slope <- apply(Slope_indices[,c(3,5,6,7,8,9,11,12, 13, 14)], 2, mean)

means <- cbind(Fjord = means_fjord, Shelf = means_shelf, Shelfbreak = means_shelfbreak, Slope= means_slope)
means <- as.data.frame(t(means))
means$Habitat <- rownames(means)
means
```
### Libraries for statistics 
```{r}
library("mgcv")
library("fields")
library("gamair")
library('car')
library("dplyr")
library("agricolae")
library("ggplot2")
library("ggthemes")
```

### Check for homogeniety of variance 
```{r}
# Levene test: P has to be over 0.05.
#leveneTest(Sp.rich ~ Habitat, center = mean, data = All_indices)
leveneTest(log(Sp.rich) ~ Habitat, center = mean, data = All_indices)#ok.
#leveneTest(sqrt(Sp.rich) ~ Habitat, center = mean, data = All_indices)
leveneTest(FRic ~ Habitat, center = mean, data = All_indices)
leveneTest(Simpson ~ Habitat, center = mean, data = All_indices)
leveneTest(FDiv ~ Habitat, center = mean, data = All_indices)
#leveneTest(Vul2 ~ Habitat, center = mean, data = All_indices)
leveneTest(Vul2^2 ~ Habitat, center = mean, data = All_indices)
leveneTest(Red ~ Habitat, center = mean, data = All_indices)
```

### ANOVA
Alongside the ANOVA test, do further checks of homogeniety of variance by looking at summary plots.
```{r}
anova1 <- aov(log(All_indices$Sp.rich) ~ All_indices$Habitat)
anova2 <- aov(All_indices$FRic ~ All_indices$Habitat)
anova3 <- aov(All_indices$Simpson ~ All_indices$Habitat)
anova4 <- aov(All_indices$FDiv ~ All_indices$Habitat)
anova5 <- aov(All_indices$Vul2^2 ~ All_indices$Habitat)
anova6 <- aov(All_indices$Red ~ All_indices$Habitat)
```
```{r}
par(mfrow=c(2,2))
plot(anova1) # sp rich --------- is a good model 
plot(anova2) # F rich  --------- is a good model 
plot(anova3) # simpson --------- is a good model 
plot(anova4) # f diversity 
plot(anova5) # vul
plot(anova6) # red
```
```{r}
summary(anova1)
summary(anova2) 
summary(anova3) # not sig.
summary(anova4)
summary(anova5)
summary(anova6)
```

### Honestly significant difference: HSD
Tukeys Mulitple comparison of treatments. This test is used to identify which groups are significantly different to one another. 
Using package 'agricolae'
```{r}
#Conduct test
hsd1 = HSD.test(aov(log(Sp.rich)~Habitat,data=All_indices), "Habitat", group=T)
hsd2 = HSD.test(aov(sqrt(FRic)~Habitat,data=All_indices), "Habitat", group=T)
hsd3 = HSD.test(aov(Simpson~Habitat,data=All_indices), "Habitat", group=T)
hsd4 = HSD.test(aov(FDiv~Habitat,data=All_indices), "Habitat", group=T)
hsd5 = HSD.test(aov(log(-All_indices$Vul2+1)~Habitat,data=All_indices), "Habitat", group=T)
hsd6 = HSD.test(aov(Red~Habitat,data=All_indices), "Habitat", group=T)
```

```{r}
# Check results
hsd1
hsd2
hsd3
hsd4
hsd5
hsd6
```

```{r}
# plot
par(mfrow=c(3,2))
plot(TukeyHSD(aov(log(Sp.rich)~Habitat,data=All_indices), "Habitat", group=T))
plot(TukeyHSD(aov(sqrt(FRic)~Habitat,data=All_indices), "Habitat", group=T))
plot(TukeyHSD(aov(Simpson~Habitat,data=All_indices), "Habitat", group=T))
plot(TukeyHSD(aov(FDiv~Habitat,data=All_indices), "Habitat", group=T))
plot(TukeyHSD(aov(log(-All_indices$Vul2+1)~Habitat,data=All_indices), "Habitat", group=T))
plot(TukeyHSD(aov(Red~Habitat,data=All_indices), "Habitat", group=T))
```

```{r}
# allocate sig. group letters
groups1 <- hsd1$groups[order(row.names(hsd1$groups)),]
groups2 <- hsd2$groups[order(row.names(hsd2$groups)),]
groups3 <- hsd3$groups[order(row.names(hsd3$groups)),]
groups4 <- hsd4$groups[order(row.names(hsd4$groups)),]
groups5 <- hsd5$groups[order(row.names(hsd5$groups)),]
groups6 <- hsd6$groups[order(row.names(hsd6$groups)),]
```

### GLM
Identifying the relationship between species and functional richness
```{r}
GLM_Fric_all <- glm(formula = All_indices$FRic ~ (log(All_indices$Sp.rich)), 
    family = (gaussian(link= log)))
summary(GLM_Fric_all)
100*(GLM_Fric_all$null.deviance-GLM_Fric_all$deviance)/GLM_Fric_all$null.deviance
# This gives the percentage explained by our model. 

chisqr <- GLM_Fric_all$null.deviance - GLM_Fric_all$deviance
GLM_Fric_all$df.null
GLM_Fric_all$df.residual
#using online chi sqr p value calculator. P=0.0001. 

# extract values you want for pl
p <- predict(GLM_Fric_all, type="response", se.fit = TRUE)
p <- data.frame(p.fit = p$fit, p.se = p$se.fit)
p <- p[order(p$p.fit),]
xpred <- data.frame(Sp.ric = (All_indices$Sp.rich))
xpred <- xpred[order(xpred$Sp.ric),]
```

## Plotting figures 
```{r}
par(mfrow=c(3,3), mar=c(4,5,2,1))
#Species & Funct richness-------------------------------------------------------
summarized = All_indices %>% group_by(Habitat) %>% summarize(Sp.rich=max(Sp.rich))
boxplot(All_indices$Sp.rich ~ All_indices$Habitat, col=brewer.pal(n = 4, name = "BrBG"),
        xlab= " ", ylab="Taxa Richness", ylim=c(0,4+(max(summarized$Sp.rich))),
        cex.axis = 1.2, cex.lab = 1.3)
title("a)", adj=0.01, line = 1, cex.main = 1.5)
points(means$Sp.rich, col = "red", pch = 18, cex=1)
over <- 3+summarized$Sp.rich
text(y=over, x=1:4, labels = groups1[,2], font = 3, cex = 1.2)
text(0.5,0.07, labels= "F = 29.0, p<0.001", col="red", font = 3, adj=0, cex = 1.2)

# Funtional Richness ------------------
summarized = All_indices %>% group_by(Habitat) %>% summarize(FRic=max(FRic))
boxplot(All_indices$FRic~All_indices$Habitat, col=brewer.pal(n = 4, name = "BrBG"),
        xlab= " ", ylab="Functional Richness", ylim = c(0,1200),
        cex.axis = 1.2, cex.lab = 1.3)
title("b)", adj=0.01, line = 1, cex.main = 1.5)
points(means$FRic, col = "red", pch = 18, cex=1)
over <- 75+summarized$FRic
text(y=over, x=1:4, labels = groups2[,2], font = 3, cex = 1.2)
text(0.5,1, labels= "F = 17.6, p<0.001", col="red", font = 3, adj=0, cex = 1.2)

#FRIC OVER SP RICH ----------------------------------------------
plot(All_indices$FRic ~ (All_indices$Sp.rich), 
     col="black", pch=19,cex.axis = 1.2, cex.lab = 1.3, cex=1.5,
     xlab=" ", ylab="Functional Richness", xlim=c(0,60), ylim=c(0,800))
grid()
points(jitter(All_indices$FRic[19:27]) ~jitter(All_indices$Sp.rich[19:27]), cex=1.5,
       col="black", bg ="darkgrey", pch=21, cex.axis = 1.2, cex.lab = 1.3)
title("c)", adj=0.01, line = 1, cex.main = 1.5)
lines(p$p.fit ~ xpred, col="black", lwd=2)
lines((p$p.fit)+(p$p.se) ~ xpred, col="grey20", lwd=1.5, lty=2)
lines(p$p.fit-p$p.se ~ xpred, col="grey20", lwd=1.5, lty=2)
text(1,750, labels= "T = 8.7, p<0.001", col="red", font = 3, adj=0, cex=1.2)

#Diversity----------------------------------------------------------------------
summarized = All_indices %>% group_by(Habitat) %>% summarize(Simpson=max(Simpson))
boxplot(All_indices$Simpson~ All_indices$Habitat, col=brewer.pal(n = 4, name = "BrBG"),
        xlab= " ", ylab="Taxa Diversity",ylim=c(0,0.1+(max(summarized$Simpson))),
        cex.axis = 1.2, cex.lab = 1.3)
title("d)", adj=0.01, line = 1, cex.main = 1.5)
points(means$Simpson, col = "red", pch = 18, cex=1)
#over <- 0.1+summarized$Simpson
#text(y=over, x=1:4, labels = groups3[,2], font = 3, cex = 0.7)
text(0.5,0.01, labels= "F = 2.0, p>0.05", col="grey40", font = 3, adj=0, cex = 1.2)


summarized = All_indices %>% group_by(Habitat) %>% summarize(FDiv=max(FDiv))
boxplot(All_indices$FDiv~All_indices$Habitat, col=brewer.pal(n = 4, name = "BrBG"),
        xlab= " ", ylab="Functional Diversity", ylim=c(0,0.1+(max(summarized$FDiv))),
        cex.axis = 1.2, cex.lab = 1.3)
title("e)", adj=0.01, line = 1, cex.main = 1.5)
points(means$FDiv, col = "red", pch = 18, cex=1)
over <- 0.05+summarized$FDiv
text(y=over, x=1:4, labels = groups4[,2], font = 3, cex = 1.2)
text(0.5,0.005, labels= "F = 3.7, p<0.05", col="red", font = 3, adj=0, cex = 1.2)


plot(All_indices$FDiv[1:18] ~ All_indices$Sp.rich[1:18], ylim=c(0.3,1.0),
     col="black", pch=19,cex.axis = 1.2, cex.lab = 1.3, cex=1.5,
     xlab=" ", ylab="Functional Diversity", xlim=c(0,60))
grid()
points(All_indices$FDiv[19:27] ~All_indices$Sp.rich[19:27],cex=1.5,
       col="black", bg ="darkgrey", pch=21, cex.axis = 1.2, cex.lab = 1.3)
title("f)", adj=0.01, line = 1, cex.main = 1.5)

#Vulnerability and Redundancy---------------------------------------------------
summarized = All_indices %>% group_by(Habitat) %>% summarize(Vul2=max(Vul2))
boxplot(All_indices$Vul2~All_indices$Habitat, col=brewer.pal(n = 4, name = "BrBG"), 
        ylab="Taxa Vulnerability", xlab = "Habitat type", ylim=c(0,1), #4+(max(summarized$Vul2))),
        cex.axis = 1.2, cex.lab = 1.3)
title("g)", adj=0.01, line = 1, cex.main = 1.5)
points(means$Vul2, col = "red", pch = 18, cex=1)
over <- 0.05+summarized$Vul2
text(y=over, x=1:4, labels = groups5[,2], font = 3, cex = 1.2)
text(0.5,0.005, labels= "F = 29.6, p<0.001", col="red", font = 3, adj=0, cex = 1.2)

summarized = All_indices %>% group_by(Habitat) %>% summarize(Red=max(Red))
boxplot(All_indices$Red ~ All_indices$Habitat, col=brewer.pal(n = 4, name = "BrBG"),
        xlab= "Habitat Type", ylab="Functional Redundancy", ylim=c(0,0.05+(max(summarized$Red))),
        cex.axis = 1.2, cex.lab = 1.3)
title("h)", adj=0.01, line = 1, cex.main = 1.5)
points(means$Red, col = "red", pch = 18, cex=1)
over <- 0.025+summarized$Red
text(y=over, x=1:4, labels = groups6[,2], font = 3, cex = 1.2)
text(0.5,0.008, labels= "F = 4.1, p<0.05", col="red", font = 3, adj=0, cex = 1.2)


plot(All_indices$Red[1:18] ~ All_indices$Sp.rich[1:18], 
     col="black", pch=19,cex.axis = 1.2, cex.lab = 1.3, cex=1.5,
     xlab="Number of Taxa", ylab="Functional Redundancy", xlim=c(0,60))
grid()
points(All_indices$Red[19:27] ~ All_indices$Sp.rich[19:27],cex=1.5,
       col="black", bg ="darkgrey", pch=21, cex.axis = 1.2, cex.lab = 1.3)
title("i)", adj=0.01, line = 1, cex.main = 1.5)
```
