---
title: "NE Greenland Trait Analysis"
author: "Phoebe Armitage"
date: '2022-April'
output:
 html_document:
  toc: TRUE
  anchor_sections: TRUE
  code_download: TRUE 
  code_folding: "hide"
  fig_caption: TRUE
  theme:
    bootswatch: "flatly"
  bibliography: references.bib
  link-citations: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
# to cite specific packages
knitr::write_bib(c("base", "bookdown", "rmarkdown"), "packages.bib")
```

### Libraries and data

To begin, we need to download our data for the analysis as well as
install any packages or libraries from R.

#### Libraries

```{r}
# load libraries 
library(vegan)     
library(devtools)  
library(ggplot2)
library(FD)
library(adiv)
library(RColorBrewer)
```

#### Read in the Data

```{r}
Taxa <- read.csv("~/Arctic Ph.D/Chapter 1. Functional analysis HMSC/Data/GitHub_NEG/Data for R/Abun_species_vegan.csv", 
                 header=TRUE, sep=",", row = 1)
Taxa <- Taxa[order(rownames(Taxa)),]
Traits <- read.csv("~/Arctic Ph.D/Chapter 1. Functional analysis HMSC/Data/GitHub_NEG/Data for R/Abun_traits_vegan.csv", 
                   header=TRUE, sep=",", row = 1)
Env <- read.csv("~/Arctic Ph.D/Chapter 1. Functional analysis HMSC/Data/GitHub_NEG/Data for R/Env.csv",  header=TRUE, sep=",", row = 1)
Env <- Env[order(rownames(Env)),]

```

### Taxa Indices

```{r}
#Species indices----------------------------------------------------------------
#Taxa indices----------------------------------------------------------------
indices <- data.frame(Abundance = rowSums(Taxa))
indices$Sp.Richness <- specnumber(Taxa)
indices$Shannon <- diversity(Taxa, index = "shannon")
indices$Simpson <- diversity(Taxa, index = "simpson")
indices$Evenness <- diversity(Taxa, index = "shannon")/log(specnumber(Taxa)) #Evenness
indices$Habitat <- as.factor(Env$Habitat)
```

#### Table with taxa indices

```{r}
knitr::kable(indices,
 caption = "Taxa Indices. Metrics calculated using Vegan package")
```

### References

@Manual{R-base, title = {R: A Language and Environment for Statistical
Computing}, author = {{R Core Team}}, organization = {R Foundation for
Statistical Computing}, address = {Vienna, Austria}, year = {2019}, url
= {<https://www.R-project.org>}, }
